---
title: "HW2 - Stat 436"
author: "Kyle Zorn"
date: "2023-03-07"
output: pdf_document
---
```{r, include=FALSE}
library(dplyr)
library(shiny)
library(tidyverse)
library(ggplot2)
```

Raw Dataset:

```{r}
winners <- read.csv("https://uwmadison.box.com/shared/static/4kj3cj1zof3qmevvn0umwoobhopz8eqk.csv")
head(winners)
```


Changes made to dataset:

1. I don't care for the runner ups and final four appearances, so I wanted to only select 'year', 'level', 'sport', 'winner', 'winner_metro'; where:
    year = year the team won
    level = college or pro level
    sport = which sport the winner was champion of
    winner = team that won
    winner_metro = the city that the team is located
    
2. I wanted to separate winner metro into two columns, 'City' and 'State' to group by City in the graphs made.

```{r}
win <- winners %>%
  select(c(year, level, sport, winner, winner_metro)) %>%
  separate(winner_metro, c('City', 'State'), sep = ",")
head(win)

```


```{r}
##define the user interface
ui <- fluidPage(
  titlePanel("Winningmost Cities (1870-2018)"),
  tabsetPanel(
    tabPanel("Scatterplot",
             selectInput("City", "Select Cities:", choices = unique(win$City), multiple = TRUE),
             plotOutput(outputId = "scatterplot", brush = brushOpts(id = "brush")),
             tableOutput("brushed_data_table")
    ),

    tabPanel("Histogram",
             sliderInput("year_range", "Select Years:", min = 1870, max = 2018, value = c(1870, 2018)),
             plotOutput(outputId = "histogram")
    )
  )
)

server <- function(input, output) {
  filtered <- reactive({
    win %>% 
      filter(City %in% input$City)
  })
  
  output$scatterplot <- renderPlot({
    ggplot(filtered()) +
      geom_point(aes(x = year, y = City, color = level)) +
      labs(title = "Winners per Year", x = "Year", y = "City")
  }, bg = "white")
  
  brushed_data <- reactive({
    brushedPoints(filtered(), input$brush, xvar = "year", yvar = "City") 
  })
    
  output$brushed_table <- renderTable({
    brushed_data()
  })
  
  output$histogram <- renderPlot({
    # Filter data based on selected years
    filtered_data <- win %>% 
      filter(year >= input$year_range[1] & year <= input$year_range[2])
    
    # Calculate the number of wins for each city
    city_wins <- filtered_data %>%
      group_by(City) %>%
      summarize(num_wins = n()) %>%
      arrange(desc(num_wins))
    
    # Select only the top 10 cities by number of wins
    top_cities <- city_wins %>% top_n(10, num_wins)
    
    # Create the histogram
    ggplot(top_cities, aes(x = reorder(City, -num_wins), y = num_wins)) +
      geom_bar(stat = "identity", fill = "steelblue") +
      labs(title = "Top 10 Winning Cities", x = "City", y = "Number of Wins") +
      theme(axis.text.x = element_text(angle = 90, hjust = 1))
  })
  
  output$brushed_data_table <- renderTable({
    brushed_data()
  })
}

shinyApp(ui, server)


```



